\documentclass[12pt, letterpaper]{article}

\date{}

\begin{document}
\noindent
\textbf{MUSIIMEMARIA EUGENIA\\
15/U/8383/EVE\\
215004147\\
Color Psychology Paper\\}

\section{Research Question:}
How does color affect one’s mood?   

\section{Literature Review:}
	
	Colors may just seem simple and unimportant, but they affect our daily lives more than we may know. If someone is feeling angry, it could just be because they’re angry, or it could be perhaps that they are surrounded by or looking at the color red. That’s right! People’s moods can change just because they are looking at different colors! There are many theories on how just a simple color can change one’s whole mood.\par
	While perceptions of color are somewhat subjective, there are some color effects that have universal meaning. Colors in the red area of the color spectrum are known as warm colors and include red, orange and yellow.
These warm colors evoke emotions ranging from feelings of warmth and comfort to feelings of anger and hostility.
Colors on the blue side of the spectrum are known as cool colors and include blue, purple and green. These colors are often described as calm, but can also call to mind feelings of sadness or indifference.\par
	According to Johnson (2007), color does affect mood by producing certain chemicals and stimulating different feelings such as hunger. For example, blue can make one feel calm because it releases calming chemicals, and red can make one hungry because it is an appetite stimulant. Yellow can make one feel irritated, and it is a fact that people lose their temper most in yellow rooms. However, pink is tranquilizing and can make one feel weak. In conclusion, Johnson says that depending on the color, one’s body can do things (like producing chemicals) that cause a certain emotional reaction (mad, sad, etc.)\par 
	Another idea, by Smith (2007), is that the effect color produces is based on what one’s body does in response. For example, yellow is mentally stimulating, and activates memory, whereas red increases confidence. Also, brown can make a person feel orderly and stable, while a dark blue can make one feel sad. Therefore, Smith says that different colors do in fact change one’s mood and the consequences can be negative or positive.\par
 	A third writer, Wollard, (2000) seems to think that color can affect one’s mood, but the effect also can depend on one’s culture and what one’s personal reflection may be. For example, someone from Japan may not associate red with anger, as people from the U.S. tend to do. Also, a person who likes the color brown may associate brown with happiness. However, Wollard does think that colors can make everyone feel the same, or close to the same, mood. According to Wollard, pink reduces aggression, which is why the walls of the jail cells in the Seattle prison are pink! Also, brown can make one feel comforted. Wollard feels that colors do affect one’s mood, but there are other factors that can alter what one is supposed to feel.\par
	 Eric, John, and Paraag’s (2007) main point about color psychology is that color has both a physiological and psychological effect. For example, green makes people feel relaxed because it relaxes their muscles and makes them breathe deeper and more slowly. Furthermore, blue lowers blood pressure, which makes one feel calm. Eric, John, and Paraag conclude that color affects one’s mood because of what it does to the body.\par
	Yet another idea, by Airey (2006), is that color is energy, and it can have a physical, mental, spiritual, and/or emotional affect on people. He states that black can make one feel sophisticated and secure, but it can also make one feel depressed. Also brown can make a person feel reliable and serious, while yellow lifts self-esteem. Therefore, Airey concludes that different colors can have different kinds of affects on people.\par


\section{Analysis:}
There are three ideas about color psychology in these sources, and they all say that color affects one’s mood. They differ based on what factors influence the effects of color, such as culture, opinion, and what goes on inside one’s body.  One of the three ideas is that color affects mood based on one’s personal opinions. For example, if a person dislikes the color pink, he may associate pink with hate. Another idea states that color affects mood based on one’s culture. For example, someone from the U.S. may think of the color green when referring to envy, while people in Japan think of yellow in connection with wanting what someone else has. However, the majority of the sources consulted say that color affects mood by influencing what goes on inside of people. For example, seeing the color blue releases calming chemicals, which in turn makes one calm. Also, because yellow is the hardest color for the eye to focus on, people may become irritated when looking at yellow, and it is a proven fact that babies cry most in yellow nurseries. \\
These theories do not seem to have much in common.


\section{Conclusion:}
Color does affect one’s mood, but it can affect boys and girls differently. For example, while most female middle school students found green neutral and balancing, most male middle school students found it secure and safe. However, there are also some similarities between the female and male middle school students. For example, most female middle school students thought that brown made them feel bored, and so did the male middle school students! Also, both found that pink made them feel love and affection. However, some results in this study were different from the research cited. For example, the research stated that brown would make girls and boys feel secure and safe, but instead it made them feel bored. Furthermore, the research saidColor does affect one’s mood, but it can affect boys and girls differently. For example, while most female middle school students found green neutral and balancing, most male middle school students found it secure and safe. However, there are also some similarities between the female and male middle school students. For example, most female middle school students thought that brown made them feel bored, and so did the male middle school students! Also, both found that pink made them feel love and affection. However, some results in this study were different from the research cited. For example, the research stated that brown would make girls and boys feel secure and safe, but instead it made them feel bored. Furthermore, the research said that yellow would make boys and girls feel irritated, but most boys and girls reported that it made them feel happy and cheery. Overall, most of the results were different from the research sources consulted. 
In any case, it is obvious that colors have a great affect on one’s mood.
 that yellow would make boys and girls feel irritated, but most boys and girls reported that it made them feel happy and cheery. Overall, most of the results were different from the research sources consulted.\\ 
In any case, it is obvious that colors have a great affect on one’s mood.
\\


\end{document}